---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

Uber and Lyft Price Prediction
Preston Chen, Megan Crawford, Mitch Feren, Alex Sampson
Prof. Huynh
Predictive Analytics Project Draft

Installing Packages
```{r}
#install.packages("tidyverse")
#install.packages("pls")
library (pls)
library(forecast)
library(zoo)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lubridate)



options(scipen = 999)
```

Loading Data, Data Exploration, Data Preparation
```{r}
# Loading in the data from a csv
cab_rides <- read.csv('Team05_Report_Data.csv')

```

```{r}
# Looking at first couple observations
head(cab_rides)
```

```{r}
# removes the rows with a duplicate ID
cab_rides <- distinct(cab_rides, cab_rides$id, .keep_all = TRUE )
cab_rides <- na.omit(cab_rides)
```

```{r}
  
# Visualizations


summary(cab_rides) #summary stats for the data


hist(cab_rides$price,xlim = c(0,110), ylim = c(0,250000), main = "Distribution of Prices", xlab = "Price in Dollars ") #distribution of prices


#Looking at the distribution of source and destination locations

counts <- table(cab_rides$source)
barplot(counts, main="Source Distribution",las = 1, horiz = TRUE, cex.names=0.3, xlim = c(0,100000))


counts2 <- table(cab_rides$destination)
barplot(counts2, main="Destination Distribution",las = 1, horiz = TRUE, cex.names=0.3, xlim = c(0,100000))



#Distribution of the type of rides 
counts3 <- table(cab_rides$name)
barplot(counts3, main="Type Distribution",las = 1, horiz = TRUE, cex.names=0.55, xlim = c(0,120000)
        ,xlab = "Frequency")


#Price by each destination

agg_dest <- aggregate(cab_rides$price, 
          by = list(cab_rides$destination), 
          FUN = mean,
          na.rm = TRUE)


agg_dest$Avg_Price <- agg_dest$x
agg_dest$Destination <- agg_dest$Group.1


ggplot(aes(x=Destination,y=Avg_Price),data=agg_dest, main = "Price by Destination")+
  geom_point() +
  scale_x_discrete(guide = guide_axis(angle = 90))


#Price by each source

agg_source <- aggregate(cab_rides$price, 
          by = list(cab_rides$source), 
          FUN = mean, 
          na.rm = TRUE)


agg_source$Avg_Price <- agg_source$x
agg_source$Source <- agg_source$Group.1


ggplot(aes(x=Source,y=Avg_Price),data=agg_source, main = "Price by Source")+
  geom_point() +
  scale_x_discrete(guide = guide_axis(angle = 90))

#Boxplot of Ride and Distance
par(mfrow=c(1,2))
boxplot(cab_rides$distance, main = "Distance")
boxplot(cab_rides$price, main = "Price")


#Boxplot of Temperature and Clouds
par(mfrow=c(1,2))
boxplot(cab_rides$temp, main = "Temperature")
boxplot(cab_rides$clouds, main = "Clouds")



#Boxplot of Pressure and Humidity
par(mfrow=c(1,2))
boxplot(cab_rides$pressure, main = "Pressure")
boxplot(cab_rides$humidity, main = "Humidity")


#Boxplot of Rain
boxplot(cab_rides$rain, main = "Rain")
```



```{r}
# Eliminating outlier rows in price column
price_outliers <- subset(cab_rides, cab_rides['price'] > 42.75)
cab_rides <- setdiff(cab_rides, price_outliers)
```

```{r}
# Setting seed and breaking data into Uber and Lyft sets
RNGkind(sample.kind = "Rounding")
set.seed(123)

uber_rides <- cab_rides %>% filter(cab_type == 1)
lyft_rides <- cab_rides %>% filter(cab_type == 0)
```

```{r}
# Separates Uber and Lyft data into training (70% of data set) and validation (30%) sets
uber.sample.rows <- sample(rownames(uber_rides), 50000)
uber_sample <- uber_rides[uber.sample.rows,]

lyft.sample.rows <- sample(rownames(lyft_rides),50000)
lyft_sample <- lyft_rides[lyft.sample.rows,]

uber.train.index <- sample(rownames(uber_sample), dim(uber_sample)[1]*0.7)
uber.train <- uber_sample[uber.train.index,]
uber.valid.index <- setdiff(rownames(uber_sample), uber.train.index)
uber.valid <- uber_sample[uber.valid.index, ]

lyft.train.index <- sample(rownames(lyft_sample), dim(lyft_sample)[1]*0.7)
lyft.train <- lyft_sample[lyft.train.index,]
lyft.valid.index <- setdiff(rownames(lyft_sample), lyft.train.index)
lyft.valid <- lyft_sample[lyft.valid.index, ]
```

Principal Component Analysis
```{r}
head(uber.train)
options(scipen = 999, digits = 2)

# Uber PCA
pca.uber <- prcomp(na.omit(uber.train[, c(3, 8, 14, 16:18, 20:21, 24:25)]), scale. = T)
summary(pca.uber)
pca.uber$rotation[,1:8]

#Lyft PCA
pca.lyft <- prcomp(na.omit(lyft.train[, c(3, 8, 14, 16:18, 20:21, 24:25)]), scale. = T)
summary(pca.lyft)
pca.lyft$rotation[,1:8]
```
Principal Component Regression
```{r}
library (forecast)
#Eliminate the non-numerical variables
uber.train.pcr <- uber.train[, c(3, 8, 14, 16:18, 20:21, 24:25)]
uber.valid.pcr <- uber.valid[, c(3, 8, 14, 16:18, 20:21, 24:25)]
lyft.train.pcr <- lyft.train[, c(3, 8, 14, 16:18, 20:21, 24:25)]
lyft.valid.pcr <- lyft.valid[, c(3, 8, 14, 16:18, 20:21, 24:25)]

#Uber PCR


library (pls)
pcr_model.uber<- pcr(price ~., data = uber.train.pcr, scale = TRUE, validation = "CV")
summary(pcr_model.uber)

pcr_predict.uber <- predict(pcr_model.uber, uber.train.pcr, ncomp = 4)

sqrt(mean((pcr_predict.uber - uber.valid.pcr$price)^2))

#Lyft PCR
```



Uber Model and Variable Selection
```{r}
# Builds out Uber model
uber.lm <- lm(price ~ source + distance + surge_multiplier + temp + clouds + name +
            pressure + rain + humidity + wind + day + hour, data = uber.train)
summary(uber.lm)
```

```{r}
# Uber model with backwards elimination
uber.lm.back <- step(uber.lm, direction = 'backward')
```

```{r}
# Model summary
summary(uber.lm.back)
# Uber: Backwards elimination model accuracy
uber.lm.back.pred <- predict(uber.lm.back, uber.valid)
accuracy(uber.lm.back.pred, uber.valid$price)
```

```{r}
# Uber model with forwards selection
uber.lm.forward <- step(uber.lm, direction = 'forward')
summary(uber.lm.forward)
```
```{r}
# Uber: Forward selection model accuracy
uber.lm.forward.pred <- predict(uber.lm.forward, uber.valid)
accuracy(uber.lm.forward.pred, uber.valid$price)
```

```{r}
# Uber model with stepwise selection
uber.lm.step <- step(uber.lm, direction = 'both')
summary(uber.lm.step)
```
```{r}
# Uber: stepwise selection model accuracy
uber.lm.step.pred <- predict(uber.lm.step, uber.valid)
accuracy(uber.lm.step.pred, uber.valid$price)
```


Lyft Model and Variable Selection
```{r}
# Builds out Lyft model
lyft.lm <- lm(price ~ source + distance + surge_multiplier + temp + clouds + name +
            pressure + rain + humidity + wind + day + hour, data = lyft.train)
summary(uber.lm)
```

```{r}
# Lyft model with backwards elimination
lyft.lm.back <- step(lyft.lm, direction = 'backward')
```


```{r}
# Model summary
summary(lyft.lm.back)
# Lyft: backwards elimination model accuracy
lyft.lm.back.pred <- predict(lyft.lm.back, lyft.valid)
accuracy(lyft.lm.back.pred, lyft.valid$price)
```

```{r}
# Lyft model with forward selection
lyft.lm.forward <- step(lyft.lm, direction = 'forward')
summary(lyft.lm.forward)
```


```{r}
# Lyft: forwards selection model accuracy
lyft.lm.forward.pred <- predict(lyft.lm.forward, lyft.valid)
accuracy(lyft.lm.forward.pred, lyft.valid$price)
```


```{r}
# Lyft model with stepwise selection
lyft.lm.step <- step(lyft.lm, direction = 'both')
summary(lyft.lm.step)
```


```{r}
# Lyft: Stepwise selection model accuracy
lyft.lm.step.pred <- predict(lyft.lm.step, lyft.valid)
accuracy(lyft.lm.step.pred, lyft.valid$price)
```

Principle Component Analysis and Principle Component Regression
```{r}

#pc11 or pc12 or none

```
```{r}
#Principal Component Regression


pcr_model<- pcr(price~., data = uber.train, scale = TRUE, validation = "CV")
summary(pcr_model)

pcr_predict <- predict(pcr_model, uber.train, ncomp = 10)

mean((pcr_predict-uber.valid)^2)


```
